Emitter Source Geolocation from Imparted Rotor Blade Modulation

### 2016-10-19 Background.tex
add bold theory figure, add legends to verify figures

The resulting modulated signal has been used in several radar applications. First to determine the specific characteristics of a noncooperative helicopter using time frequency analysis in \cite{Bullard1991} that defines a specific signature for physical parameters of a helicopter. The ability to detect a hovering helicopter based on the blade flashes caused by reflected radar pulses off the blades is examined in \cite{Misiurewicz1997} where they used information the time domain to determine if a helicopter was hovering. The micro-Doppler produced by the radar reflection in \cite{Thayaparan2007} is used to determine the rotation rate of both the main and tail rotor blades using the wavelet transform. Other types of aircraft produce modulations on a returned radar pulse, one type known as Jet Engine Modulation (JEM) is the product of a radar pulse reflecting off the jet engine blades \cite{NavalAirSystemsCommand1999}. The information can similarly be used to quantify physical aspects of the aircraft when present in order to perform classification.

The Doppler shifted signal caused by RBM is shown in \cite{F.BarbieroF.VincentT.Deloues2014} to effect the performance of common adaptive Global Positioning System (GPS) anti-jamming algorithms located on helicopter platforms. They were able to mitigate the effects by estimating the covariance within a few milliseconds but this comes at an extreme computation overhead. And for iterative based anti-jamming approaches the non-stationary signal characteristics cause degradation as well. The affect of RBM on the Bit Error Rate (BER) of communications is examined in \cite{YiminZhang2000} where effects of multipath and Doppler cause large BER values at small elevation angles and large scattering coefficients described in the model.

@@

%describe the situation and the position of the objects
There are three physical objects within this scene. The first being the rotor blade, which is the dark green bar depicted in the figure. Next is the transmitter $Tx$, in green, and the receiver $Rx$ in blue. The measurements, depicted as black bars with denoting start and stop points, reference the various distances used in the following equations. The dashed green line shows at which altitude the helicopter is hovering at, and the red dashed line transverses the transmitter and the center of rotation for the blades. $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades and $\theta_r$ denotes the specular reflection of the casted and reflected rays in orange. Lastly the line of sight (LOS) ray shows that the transmitter and receiver are in visual line of sight of each other.

@@

where $f_d$ is the Doppler frequency shift, $f_{DopplerMax}$ is described by equation \ref{eqn:fd_max} as the maximum amount of Doppler shift possible, $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades, and $\theta_{b}$ is the current angle of the rotor blade.

@@

%validation consideration
Since this model has the receiver located at the axis of rotation this model can be seen as a special case of the next 3D models, this is because it is symmetric around the axis of rotation. Therefore, any placement of the transmitter in 3D space can be distilled down to a 2D profile as such. This symmetry should then be mirrored in the ray tracing simulation as verification.

@@

%describe the situation and the position of the objects
The objects in \ref{fig:3D_model_90az}  are similar to the ones previously stated in the 2D case. The X and Y axis are specifically drawn in blue and red respectfully, with the Z axis now the green dashed line and altitude is now denoted with a measurement bar. The receiver offset in the +Y direction is now stated. and the Blades are now colored gray. All other variables and colors denote their objects as previously stated in the 2D case.

@@

where $MaxUpperEnvelope$ is an array of the maximum upper envelope values for each elevation sample, $MinLowerEnvelope$ is an array of minimum lower envelope values for each elevation sample, and $\theta_{Limited}$ is the limited elevation angle range due to

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{([0,\dots, (BladeLength - Rx_y)])}\right)
	\label{eq:theory_90_limited}
\end{equation}

@@

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{([0, \dots, BladeLength)}\right)
	\label{eq:theory_270_limited}
\end{equation}

@@

and the MinLowerEnvelope is set to 0 for the range.

@@

\subsubsection{Receding Model}
%go over the 3d model for the receding blade
The receding model shown in Figure \ref{fig:3D_model_0az} depicts the transmitter at an azimuth of 0\textdegree \space where the axis positions back to their original places and the direction of the rotor blade is now depicted in red.

@@

%formulate the geometric equations
\begin{equation}
	MinLowerEnvelope = f_{DopplerMax}cos(\theta_{El})
	\label{eq:theory_0_lower}
\end{equation}

and the MaxUpperEnvelope is set to 0 for the range.

>>>>

The resulting modulated signal has been used in several radar applications. One radar application of RBM was to determine the specific characteristics of a noncooperative helicopter using time frequency analysis by Bullard and Dowdy in \cite{Bullard1991}. Where they define specific signatures for physical parameters of a helicopter. The ability to detect a hovering helicopter based on the blade flashes caused by reflected radar pulses off the blades is examined by Misiurewicz, Kulpa, and Czekala in \cite{Misiurewicz1997}, where they used information the time domain to determine if a helicopter was hovering. The micro-Doppler produced by the radar reflection was used by Thayaparan et. al. in \cite{Thayaparan2007} to determine the rotation rate of both the main and tail rotor blades using the Wavelet transform. Other types of aircraft produce modulations on a returned radar pulse; one type known as Jet Engine Modulation (JEM) is the product of a radar pulse reflecting off the jet engine blades \cite{NavalAirSystemsCommand1999}. The information can similarly be used to quantify physical aspects of the aircraft when present in order to perform classification.

The Doppler shifted signal, caused by RBM, is shown by Barbiero, Vincent, and Deloues in \cite{F.BarbieroF.VincentT.Deloues2014} to effect the performance of common adaptive Global Positioning System (GPS) anti-jamming algorithms located on helicopter platforms. They were able to mitigate the effects by estimating the covariance within a few milliseconds but this comes at an extreme computation overhead. And for iterative based anti-jamming approaches the non-stationary signal characteristics cause degradation as well. The affect of RBM on the Bit Error Rate (BER) of communications is examined by Zhang, Amin, and Mancuso in \cite{YiminZhang2000}, where effects of multipath and Doppler cause large BER values at small elevation angles and large scattering coefficients described in the model.

@@

%describe the situation and the position of the objects
There are three physical objects within this scene. The first being the rotor blade, which is the dark green bar depicted in the figure. Next is the transmitter $Tx$, in green, and the receiver $Rx$ in blue. The measurements, depicted as black bars with denoted start and stop points, reference the various distances used in the following equations. The dashed green line shows at which altitude the helicopter is hovering at, and the red dashed line transverses the transmitter and the center of rotation for the blades. $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades and $\theta_r$ denotes the specular reflection of the casted and reflected rays in orange. Lastly the line of sight (LOS) ray shows that the transmitter and receiver are in visual line of sight of each other.

The resulting Doppler equation is described by

@@

where $f_d$ is the Doppler frequency shift, $f_{DopplerMax}$ is defined later by equation \ref{eqn:fd_max} as the maximum amount of Doppler shift possible, $\theta_{El}$ is the elevation angle of the transmitter with respect to the rotor blades, and $\theta_{b}$ is the current angle of the rotor blade.

@@

%validation consideration
Since this model has the receiver located at the axis of rotation this model can be seen as a special case of the next 3D models, because it is symmetric around the axis of rotation. Therefore, any placement of the transmitter in 3D space can be distilled down to a 2D profile as such. This symmetry should then be mirrored in the ray tracing simulation as verification.

@@

%describe the situation and the position of the objects
The objects in \ref{fig:3D_model_90az}  are similar to the ones previously stated in the 2D case. The X and Y axis are specifically drawn in blue and red respectfully, with the Z axis now the green dashed line and altitude is now denoted with a measurement bar. The receiver offset in the +Y direction is now stated and the blades are now colored gray. All other variables and colors denote their objects as previously stated in the 2D case.

@@

where $MaxUpperEnvelope$ is an array of the maximum upper envelope values for each elevation sample, $MinLowerEnvelope$ is an array of minimum lower envelope values for each elevation sample, and $\theta_{Limited}$ is the limited elevation angle range

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{[0,\dots, (BladeLength - Rx_y)]}\right)
	\label{eq:theory_90_limited}
\end{equation}

@@

\begin{equation}
	\theta_{Limited} = atan\left(\frac{Rx_z}{[0, \dots, BladeLength]}\right)
	\label{eq:theory_270_limited}
\end{equation}

@@

and the $MinLowerEnvelope$ is set to 0 for the range.

@@

\subsubsection{Receding Model}
%go over the 3d model for the receding blade
The receding model shown in Figure \ref{fig:3D_model_0az} depicts the transmitter at an azimuth of 0\textdegree \space where the axis positions are back to their original places and the direction of the rotor blade is now depicted in red.

@@

%formulate the geometric equations
\begin{equation}
	MinLowerEnvelope = -f_{DopplerMax}cos(\theta_{El})
	\label{eq:theory_0_lower}
\end{equation}

and the $MaxUpperEnvelope$ is set to 0 for the range.

### 2016-10-19 Conclusion.tex
add bold theory figure, add legends to verify figures

Over the course of this thesis a three dimensional ray tracer application was built in C++ to produce a RBM signal according to a variety of physical parameters. The physical parameters associated with a helicopter blade and the positioning of transmitter and receiver hardware were swept independently to build knowledge on how they affect the resulting signal. That data was then processed using time frequency analysis to produce a doppler envelope. An envelope of each parameter step was used to determine an overall trend. Then with those trends an algorithm was derived to find the azimuth and elevation angle of the transmitter with relationship to the helicopter platform. The algorithm accurately finds the azimuth angle and elevation using the data collected in the simulation environment with almost no error in the azimuth estimation and less than 2\% error in elevation angle when blade pitch in 0\textdegree.

\section{Future Work}
%Real world experiment needs to be conducted starting with small scale and moving towards a large scale implementation on a helicopter platform.
The first thing that needs to be done in continuing this research is to perform a real world experiment to officially evaluate the ray tracing application. To do this appropriate hardware needs to be defined in the form of a software defined radio for both the receiver and transmitter along with a software workflow such as GNU Radio. Next a small scale environment would need to be set up, with some form of rotating metal blade or thin plate rotating at a sufficient rate to produce a doppler spread similar to those defined in the simulation. After that data is collected and analyzed a full scale experiment would need to be conducted including a real helicopter platform and an environment with large expanses of flat ground.

@@

%Because there is more information in the received envelopes besides the maximum and minimum the specific shape could be used to better estimate azimuth and elevation without the need to complete a full revolution as described in the proposed method.
The estimation methods described in this thesis rely only on the maximum and minimum values of the doppler envelopes calculated but their shape could also be of particular interest as well. When varying the parameters of azimuth and elevation the shape of the envelope changes in a distinctive manner, by using a training algorithm the position of the transmitter might be able to be tracked in time by evaluating the change in shape. Under the defined estimation methods a full revolution of the helicopter platform is needed to resolve position ambiguity.

>>>>

Over the course of this thesis a three dimensional ray tracer application was built in C++ to produce a RBM signal according to a variety of physical parameters. The physical parameters associated with a helicopter blade and the positioning of transmitter and receiver hardware were swept independently to build knowledge on how they affect the resulting signal. That data was then processed using time frequency analysis to produce a Doppler envelope. An envelope of each parameter step was used to determine an overall trend. Then with those trends an algorithm was derived to find the azimuth and elevation angle of the transmitter with relationship to the helicopter platform. The algorithm accurately finds the azimuth angle and elevation using the data collected in the simulation environment with almost no error in the azimuth estimation and less than 2\% error in elevation angle when blade pitch in 0\textdegree.

\section{Future Work}
%Real world experiment needs to be conducted starting with small scale and moving towards a large scale implementation on a helicopter platform.
The first thing that needs to be done in continuing this research is to perform a real world experiment to officially evaluate the ray tracing application. To do this appropriate hardware needs to be defined in the form of a software defined radio for both the receiver and transmitter along with a software workflow such as GNU Radio. Next a small scale environment would need to be set up, with some form of rotating metal blade or thin plate rotating at a sufficient rate to produce a Doppler spread similar to those defined in the simulation. After that data is collected and analyzed a full scale experiment would need to be conducted including a real helicopter platform and an environment with large expanses of flat ground.

@@

%Because there is more information in the received envelopes besides the maximum and minimum the specific shape could be used to better estimate azimuth and elevation without the need to complete a full revolution as described in the proposed method.
The estimation methods described in this thesis rely only on the maximum and minimum values of the Doppler envelopes calculated but their shape could also be of particular interest as well. When varying the parameters of azimuth and elevation the shape of the envelope changes in a distinctive manner, by using a training algorithm the position of the transmitter might be able to be tracked in time by evaluating the change in shape. Under the defined estimation methods a full revolution of the helicopter platform is needed to resolve position ambiguity.

### 2016-10-19 Introduction.tex
add bold theory figure, add legends to verify figures

\section{Organization}
In Chapter \ref{ch:background} we present the background and theory behind rotor blade modulation. we then formulate the basic geometry behind the RBM effect.

In Chapter \ref{ch:radio_propagation} we Introduce the radio propagation simulation in several sections. First is the RF propagation model which defines the physical effects that are simulated within the software. Next the objects of the scene model, as well as the dimensionality, are defined in software to be representative of their physical  dimensions and functionality. Lastly, we describe the core ray tracing engine that traces the scene in successive frames to build the receiver output.

@@

In Chapter  \ref{ch:results} Those characteristics are then used to estimate the azimuth and elevation angle of the transmitter based on the given rotor and receiver parameters. Those estimates are then compared with the original position of the transmitter to determine the methods accuracy.

>>>>

\section{Organization}
In Chapter \ref{ch:background} we present the background and theory behind rotor blade modulation. We then formulate the basic geometry behind the RBM effect.

In Chapter \ref{ch:radio_propagation} we introduce the radio propagation simulation in several sections. First is the RF propagation model which defines the physical effects that are simulated within the software. Next the objects of the scene model, as well as the dimensionality, are defined in software to be representative of their physical  dimensions and functionality. Lastly, we describe the core ray tracing engine that traces the scene in successive frames to build the receiver output.

@@

In Chapter  \ref{ch:results} those characteristics from chapter \ref{ch:simulations} are then used to estimate the azimuth and elevation angle of the transmitter based on the given rotor and receiver parameters. Those estimates are then compared with the original position of the transmitter to determine the methods accuracy.

### 2016-10-19 MSc. Thesis Template.tex
add bold theory figure, add legends to verify figures

% The Acknowledgements page, for thanking everyone
\acknowledgements{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
I would like to thank my thesis advisor Dr. Tamal Bose for the opportunity to work in his lab, and for the guidance through graduate school. I would also like to thank Matt Kruse and Charlie Cooper for coming up the research topic and support through the process. I would also like to thank the Broadband Wireless Access \& Applications Center and Rincon Research Corporation for supporting this research.
}
\clearpage  
% End of the Acknowledgements

@@

\setstretch{1.3}  % Return the line spacing back to 1.3
\pagestyle{empty}  % Page style needs to be empty for this page
\dedicatory{To my Mother and Father, \\my brother Danny, \\and my lovely girlfriend Clarice,\\all with love, \\Thank you for your unconditional support.}
\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics

@@

%% ----------------------------------------------------------------
\setstretch{1.5}  % Set the line spacing to 1.5, this makes the following tables easier to read
%\clearpage  % Start a new page
\lhead{\emph{Abbreviations}}  % Set the left side page header to "Abbreviations"
\listofsymbols{ll}  % Include a list of Abbreviations (a table of two columns)
{
 %\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
 \textbf{2D} & 2-\textbf{D}imensional \\
 \textbf{3D} & 3-\textbf{D}imensional \\
 \textbf{BER} & \textbf{B}it \textbf{E}error \textbf{R}ate \\
 \textbf{BRDF} & \textbf{B}i-directional \textbf{R}eflection  \textbf{D}istrobution \textbf{F}unction \\
 \textbf{CSV} & \textbf{C}omma \textbf{S}eparated \textbf{V}alue \\
 \textbf{DSP} & \textbf{D}igital \textbf{S}ignal \textbf{P}rocessing \\
 \textbf{GPS} & \textbf{G}lobal \textbf{P}ositioning \textbf{S}ystem \\
 \textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit \\
 \textbf{GUI} & \textbf{G}raphical \textbf{U}ser \textbf{I}nterface \\
 \textbf{IQ} & \textbf{I}n-phase and \textbf{Q}uadrature \\
 \textbf{JEM} & \textbf{J}et \textbf{E}ngine \textbf{M}odulation \\
 \textbf{LOS} & \textbf{L}ine \textbf{O}f \textbf{S}ight \\
 \textbf{RBM} & \textbf{R}otor \textbf{B}lade \textbf{M}odulation \\
 \textbf{RF} & \textbf{R}adio \textbf{F}requency \\
 \textbf{RPM} & \textbf{R}otations \textbf{P}er \textbf{M}inute \\
 \textbf{Rx} & \textbf{R}eceiver \\
 \textbf{STFT} & \textbf{S}hort \textbf{T}ime \textbf{F}ourier \textbf{T}ransform \\ 
 \textbf{Tx} & \textbf{T}ransmitter \\

@@

%% ----------------------------------------------------------------
%\clearpage  % Start a new page
%\lhead{\emph{Mathematical Notations}}  % Set the left side page header to "Physical Constants"
%\listofconstants{ll}  % Include a list of Physical Constants (a four column table)
{
% Constant Name & Symbol & = & Constant Value (with units) \\

%% ----------------------------------------------------------------
%\clearpage  %Start a new page
%\lhead{\emph{Symbols}}  % Set the left side page header to "Symbols"
%\listofnomenclature{ll}  % Include a list of Symbols (a three column table)
{
% symbol & name & unit \\

}
%% ----------------------------------------------------------------
% End of the pre-able, contents and lists of things
% The Abstract Page
\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
\abstract{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
In RF communications with a rotorcraft, such as a helicopter, the rotor blades can impart a modulation onto the received signal called Rotor Blade Modulation (RBM). This modulation is caused by the reflection of a signal off the rotating blades. The reflected signal is Doppler shifted based on where the signal is reflected along the length of the blade as well as the angle between the axis of rotation and the emitter. RBM is known to degrade the performance of RF communications on rotorcraft and can be used in radar applications to detect and classify aircraft, but there is little on its usefulness in other areas. This thesis looks at the ability to utilize the RBM phenomenon on the rotorcraft itself to geo-locate and track a signal emitter on the ground.

To do this a 3D RF ray tracing program was developed in C++ to produce simulations of RBM signals. The developed program is based on optical ray tracing algorithms with modified physical propagation effects for RF signals, and swapping lights and cameras for RF transmitters and receivers respectively. The ray tracer was then run over a realistic set of physical parameters to determine their effects on the received signal, this includes transmitter azimuth and elevation angle, receiver position, blade pitch, etc. along with their combinations. From the simulations the azimuth and elevation angle, of the transmitter, produce predictable modulations on the received signal.  Based on the trends in the signal's modulation a DSP algorithm was distilled down that accurately determines the azimuth and elevation angle of the transmitter from simulated signal data.
}

>>>>

% The Acknowledgements page, for thanking everyone
\acknowledgements{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
I would like to thank my thesis advisor Dr. Tamal Bose for the opportunity to work in his lab and for the guidance through graduate school. I would also like to thank Matt Kruse and Charlie Cooper for coming up the research topic and support through the process. I would also like to thank the Broadband Wireless Access \& Applications Center and Rincon Research Corporation for supporting this research.
}
\clearpage  
% End of the Acknowledgements

@@

\setstretch{1.3}  % Return the line spacing back to 1.3
\pagestyle{empty}  % Page style needs to be empty for this page
\dedicatory{To my Mother and Father, \\my brother Danny, \\and my lovely girlfriend Clarice,\\all with love, \\thank you for your unconditional support.}
\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics

@@

%% ----------------------------------------------------------------
\setstretch{1.5}  % Set the line spacing to 1.5, this makes the following tables easier to read
\clearpage  % Start a new page
\lhead{\emph{Abbreviations}}  % Set the left side page header to "Abbreviations"
\listofsymbols{ll}  % Include a list of Abbreviations (a table of two columns)
{
 %\textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
 \textbf{2D} & 2-\textbf{D}imensional \\
 \textbf{3D} & 3-\textbf{D}imensional \\
 \textbf{BER} & \textbf{B}it \textbf{E}rror \textbf{R}ate \\
 \textbf{BRDF} & \textbf{B}i-directional \textbf{R}eflection  \textbf{D}istribution \textbf{F}unction \\
 \textbf{CSV} & \textbf{C}omma \textbf{S}eparated \textbf{V}alue \\
 \textbf{DSP} & \textbf{D}igital \textbf{S}ignal \textbf{P}rocessing \\
 \textbf{GPS} & \textbf{G}lobal \textbf{P}ositioning \textbf{S}ystem \\
 \textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit \\
 \textbf{GUI} & \textbf{G}raphical \textbf{U}ser \textbf{I}nterface \\
 \textbf{IQ} & \textbf{I}n-phase and \textbf{Q}uadrature \\
 \textbf{JEM} & \textbf{J}et \textbf{E}ngine \textbf{M}odulation \\
 \textbf{LOS} & \textbf{L}ine \textbf{O}f \textbf{S}ight \\
 \textbf{RBM} & \textbf{R}otor \textbf{B}lade \textbf{M}odulation \\
 \textbf{RF} & \textbf{R}adio \textbf{F}requency \\
 \textbf{RPM} & \textbf{R}otations \textbf{P}er \textbf{M}inute \\
 \textbf{Rx} & \textbf{R}eceiver \\
 \textbf{STFT} & \textbf{S}hort \textbf{T}ime \textbf{F}ourier \textbf{T}ransform \\ 
 \textbf{Tx} & \textbf{T}ransmitter \\

@@

%% ----------------------------------------------------------------
%\clearpage  % Start a new page
%\lhead{\emph{Mathematical Notations}}  % Set the left side page header to "Physical Constants"
%\listofconstants{ll}  % Include a list of Physical Constants (a four column table)
%{
% Constant Name & Symbol & = & Constant Value (with units) \\

%}

%% ----------------------------------------------------------------
\clearpage  %Start a new page
\lhead{\emph{Symbols}}  % Set the left side page header to "Symbols"
\listofnomenclature{l p{5cm}| l}  % Include a list of Symbols (a three column table)
{
 \textbf{Symbol} & \textbf{Name} & \textbf{Units} \\
 $\theta_{El}$ & Transmitter Elevation angle & degrees \\
 $\theta_{r}$ & Ray Reflection angle & degrees \\
 $\theta_{Az}$ & Transmitter Azimuth angle & degrees \\
 $f_d$ & Doppler Frequency shift & Hz \\
 $\theta_b$ & Rotor Blade angle & degrees \\
 $Rx_x$ & Receiver position in X axis & meters \\
 $Rx_y$ & Receiver position in Y axis & meters \\
 $Rx_z$ & Receiver position in Z axis & meters \\
 $G_r$ & Receiver antenna gain & \\
 $G_t$ & Transmitter antenna gain& \\
 $R$ & Distance & meters\\
 $\vec{n}$ & Surface normal vector & \\
 $\vec{d_i}$ & Direction vector of incident ray & \\
 $\Gamma$ & Reflection loss coefficient & \\
 $\theta_i$ & Angle of incidence with an objects normal & radians \\
 $d$ & Total distance traveled by a ray & meters \\ 
 $t_0$ & Starting time in a frame & seconds \\
 $r$ & A radius along the length of the blade & meters \\
 $\omega_r$ & Rotational velocity of the blade & rad/second \\
 $F$ & Fading coefficient & \\
 $\lambda$ & Wavelength & meters \\
 $c$ & Speed of Light & meters/seconds \\
 $f$ & frequency & Hz \\
 $f_0$ & initial frequency & Hz \\
 $f_s$ & sampling frequency & samples/seconds \\
 $EnvelopeDifference$ & difference between two doppler envelopes & Hz \\
 $MaxUpperEnvelope$ & Array of values corresponding to the maximum of each azimuth or elevation upper envelope calculation & Hz \\
 $MinLowerEnvelope$ & Array of values corresponding to the minimum of each azimuth or elevation lower envelope calculation & Hz \\
 $BladeLength$ & Length of single rotor blade (radius of rotor) & meters \\
 $f_c$ & Center Frequency & Hz \\
 $R_{Estimate}$ & Estimated reflection radius & meters \\
 $f_{DopplerMax}$ & The maximum amount of doppler for a particular blade length and RPM & Hz \\
 $f_{DopplerCorrect}$ & The corrected doppler per $\theta_{Az}$ to be processed & Hz \\
 $A_{shift}$ & The amount of shift needed to correct for Blade pitch & Hz \\
}
%% ----------------------------------------------------------------
% End of the pre-able, contents and lists of things
% The Abstract Page
\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
\abstract{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
In RF communications with a rotorcraft such as a helicopter, the rotor blades can impart a modulation onto the received signal called Rotor Blade Modulation (RBM). This modulation is caused by the reflection of a signal off the rotating blades. The reflected signal is Doppler shifted based on where the signal is reflected along the length of the blade as well as the angle between the axis of rotation and the emitter. RBM is known to degrade the performance of RF communications on rotorcraft and can be used in radar applications to detect and classify aircraft, but there is little on its usefulness in other areas. This thesis looks at the ability to utilize the RBM phenomenon on the rotorcraft itself to geo-locate and track a signal emitter on the ground.

To do this a 3D RF ray tracing program was developed in C++ to produce simulations of RBM signals. The developed program is based on optical ray tracing algorithms with modified physical propagation effects for RF signals, and swapping lights and cameras for RF transmitters and receivers respectively. The ray tracer was then run over a realistic set of physical parameters to determine their effects on the received signal; this includes transmitter azimuth and elevation angle, receiver position, blade pitch, etc. along with their combinations. The simulations of the azimuth and elevation angle produce predictable modulations on the received signal. Based on the trends in the signal's modulation, a DSP algorithm was distilled down that accurately determines the azimuth and elevation angle of the transmitter from simulated signal data.
}
